---
title: "WildYeastReport: AlcoholToleranceStudy"
date: "`r format(Sys.Date())`"
author:
- Carlie Barnhill^1^*
- Jacob MacPhee^1^*
- Kenneth McLaughlin^1^*

output:
  bookdown::html_document2:
    code_folding: show
    keep_md: yes
    toc: TRUE
    toc_float: TRUE
    toc_depth: 6
    fig_caption: yes
editor_options: 
  markdown: 
    wrap: 72
---
```{css, echo=FALSE}
p.caption {
  font-size: 18px;
}
```

# Affiliations {-}
^1^Mount Allison University, New Brunswick, Canada  

# Acknowledgements {-}
^1^Maximilian Berthold and Douglas A. Campbell


This .Rmd imports Molecular Device Absorbance data, with data reorganized into columns labelled by treatment.

```{r load libraries, echo=FALSE, message = FALSE, warning = FALSE} 
# libraries; Note check actual dependencies
library(tidyverse) #core tidyverse packages
library(growthrates)
library(googlesheets4)
library(growthrates)
library(broom)
library(kableExtra)
```


# Settings for file import

# Note: The .tsv file exported from the Molecular Dynamics software with encoding UTF16LE had embedded null characters that caused problems upon import.
As a hacked solution, download the .tsv from 'Teams'.
Import the .tsv into a new GoogleSheet.
Delete the first two rows of the file.
Export from GoogleSheet as .csv.
Move the .csv to the .Rproj folder.
A better way would be code to read the problematic UTF16LE file with null characters.
The issue relates to a complex file structure exported from Molecular Dynamics software.
```{r variable names for file import & processing MOLECULAR DEVICES}
Project <- "WildYeast"

#set variables for file import & processing
DataPath <- file.path("..", "Data", "RawData", fsep = .Platform$file.sep)
file_id <- ".csv"

# DataOneDrive <- "~/OneDrive - Mount Allison University/BIOL2201_2024/StudentDataTest"

```


```{r Retrieve ATMetaData}
# gs4_auth()

#Instead of sending a token, googlesheets4 will send an API key. This can be used to access public resources for which no Google sign-in is required. 
googlesheets4::gs4_deauth()

# Define the Google Sheet URL (replace with your actual URL)
sheet_url <- "https://docs.google.com/spreadsheets/d/17dDzASxhWDbVpQFXb201vT2oB0rkyi2h4gG33ArOaDA/edit?usp=sharing"

# Read the Google Sheet into R as a data frame
ATMetaData <- read_sheet(sheet_url)

# View the imported data (optional)
# View(ATMetaData)
```

# List Data file(s)
```{r datafiles}
DataFiles <- list.files(path = DataPath, pattern = file_id, full.names = TRUE)

DataFiles

FileEncodeMD <- as.character(guess_encoding(file = DataFiles[1])[1,1])
FileEncodeMD
```

# Read ODData file(s) into form readable by R
```{r dataread}
ODData <- readr::read_csv(file =  DataFiles[1])

head(ODData)


#data.table::fread(text = readLines(DataFiles[1], skipNul = T))


```


```{r read all files}
#code for reading multiple files, if needed
# read_delim_plus <- function(flnm, delimiter, headerrows, fileencode){read_delim(flnm, delim = delimiter,  col_names = TRUE,  skip = headerrows, escape_double = FALSE,  locale = locale(encoding = fileencode), trim_ws = TRUE) |>
#     mutate(Filename = flnm)
#   }
# 
# ODData <- DataFiles |>
#   map_df(~read_delim_plus(flnm = ., delimiter = DelimiterMD,  headerrows = HeaderRowsMD,  fileencode = FileEncodeMD))
# 
# head(ODData)
```

# Molecular Dynamics exports problematic variable names; fix them
```{r variable names}
colnames(ODData)
```

```{r rename varible}
# Rename column using pattern matching
ODData <- ODData %>%
  rename_with(~ gsub("Temperature.*", "TempC", .x))  # Replace any name starting with "Temperature"
```


```{r verify rename worked}
head(ODData)
```

# Lost labels for OD680 & OD720 in hacked import; regenerate them.

```{r OD labels}

#remove blank rows by filtering out rows where Temp_C is NA
ODData <- ODData |>
  filter(!is.na(TempC))

#this is a hack solution, knowing there are only 15 rows of data
#better way would be to keep the OD settings during import
ODData <- ODData |>
  mutate(nm = ifelse(row_number() %in% c(1:15), 680, 720), .before = Time)

head(ODData)
```


# For plotting it is easier to have data in 'long' format, where Well becomes a Value of a Variable, rather than a separate variable for each well.
```{r ODData long}
ODDataLong <- pivot_longer(ODData, cols = -c(nm, Time, TempC), names_to = "Well", values_to = "OD")

ODDataLong |>
  ggplot() +
  geom_point(aes(x = Time, y = OD)) +
  facet_grid(cols = vars(nm))

head(ODDataLong)
```


# Data in long format (rds).
```{r save rds}
saveRDS(ODDataLong, file = file.path("..", "Data", "ProcessedData", "ODDataLong.rds"))
```

# Convert time from h:m:s format to numeric format (1,2,3,4,5,6,7,8 etc.)

```{r}
ODDataLong <- ODDataLong %>% 
  mutate(Time_numeric = as.numeric(hms(Time)) / 3600)  # Converts to hours
```

# Merge ODDataLong with Metadata before attempting data analyses.

```{r Merge ODData with Metadata}
# Merge ODData with ATMetaData
Merged_ATData <- ODDataLong %>%
  left_join(ATMetaData, by = "Well")
#View(Merged_ATData)
```

# Filter out wells that had no media/innoculant to avoid problems in analysis.
```{r Filter NA Results}

# Exclude wells G10, G11, G12, H10, H11, H12
Merged_ATData <- Merged_ATData %>%
  filter(!Well %in% c("G10", "G11", "G12", "H10", "H11", "H12"))

# Verify
table(Merged_ATData$Well) 
```
# Filter merged data by wavelength, inorder to make  growth curves for OD720 and OD680.
```{r Filter Data by nm}
# Filter data for OD 680 nm
Merged_ATData_680 <- Merged_ATData %>%
  filter(nm == 680)

# Filter data for OD 720 nm
Merged_ATData_720 <- Merged_ATData %>%
  filter(nm == 720)
```

# Double check data is in time numeric form for analyses.

```{r Convert to time numeric}
Merged_ATData <- Merged_ATData %>% 
  mutate(Time_hour = as.numeric(hms(Time)) / 3600)  # Converts to hours
```

# Verify structure prior to spline test for MuMax, problematic structure, will cause problems.

```{r Verify Merged_ATData Structure}
#Verify dataset structure
str(Merged_ATData)  
head(Merged_ATData) 
```
# Growth Curves at OD720.
```{r Growth Curves at OD720}

Merged_ATData |>
  filter(nm == 720)|>
ggplot(aes(x = Time_hour, y = OD, color = as.factor(SampleID), group = Well)) +
  geom_line() +
  geom_point() +
  facet_grid(rows = vars(EthanolConcentration)) + 
#  labs(
 #   title = "Growth Curves for Each Sample (OD at 720 nm)",
  #  x = "Time (hours)", 
   # y = "Optical Density (OD720)",
    #color = "Ethanol (%)"
  #) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    legend.position = "bottom"  
  )
```
# Growth Curves at OD680.
```{r Growth Curves at OD680}
Merged_ATData |>
  filter(nm == 680)|>
ggplot(aes(x = Time_numeric, y = OD, color = as.factor(SampleID), group = Well)) +
  geom_line() +
  geom_point() +
  facet_grid(rows = vars(EthanolConcentration)) + 
#  labs(
 #   title = "Growth Curves for Each Sample (OD at 720 nm)",
  #  x = "Time (hours)", 
   # y = "Optical Density (OD720)",
    #color = "Ethanol (%)"
  #) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    legend.position = "bottom"  
  )
```

# Spline Fitting for Determining MuMax and R-squared for all Wells.

```{r Run splines test for MuMax and R-squared}

# Step 2: Spline Fitting and Parameter Extraction
SplineRates_nest <- Merged_ATData |> 
  group_by(nm, Well, Source, SampleID, EthanolConcentration, BrothType, GrowthTempC) |> 
  nest() |> 
  mutate(SplineFit = purrr::map(data, ~tryCatch(
      growthrates::fit_spline(.$Time_hour, .$OD),  # Fit spline
      error = function(e) NULL  # Handle errors gracefully
    )), # Fit spline with error handling
Mumax_hour = purrr::map_dbl(SplineFit, ~if (!is.null(.)) pluck(., "par", "mumax") else NA_real_),  
    rsquared = purrr::map_dbl(SplineFit, ~if (!is.null(.)) pluck(., "rsquared") else NA_real_))|> 
  ungroup()

SplineRatesResults <- SplineRates_nest |> 
  select(nm, Well, Mumax_hour, rsquared, Source, SampleID, EthanolConcentration, BrothType, GrowthTempC)

# View the results
print(SplineRatesResults)

```


# Cleaned the resulting merged data, "Merged_ATMetaData_MuMax", before plotting MuMax vs EtOH concentration.

```{r Filter NA values}
# Remove rows with any NA values
SplineRatesResults <- na.omit(SplineRatesResults)

# Check the result
head(SplineRatesResults)

#View(SplineRatesResults)

```

# Fit Linear Models
```{r}
#define exponential decay function for data fitting.
#exp_decay <- function(x, i, mu){y = i * exp(mu * x)}

#need to merge left_join(Merged_ATMetaData_MuMax_clean, true MetaData so you can talk about growth responses to EtOH in a sensible manner
 # left_join(Merged_ATMetaData_MuMax_clean, 

MuEtOH_nest <- SplineRatesResults |>
  nest(.by = c("SampleID", "nm", "Source", "BrothType", "GrowthTempC")) |> #want to nest by source
  mutate(LinearFit = purrr::map(data, ~lm(Mumax_hour ~ EthanolConcentration,
                                            data = .x)),
         LinearTidy = purrr::map(LinearFit, tidy),
         LinearParam = purrr::map(LinearFit, glance),
         LinearPredict = purrr::map(LinearFit, augment))
```

```{r plot mumax vs EtOh}
temp_labeller <- function(value) {
  paste(value, "°C")
}
nm_labeller <- function(value) {
  paste(value, "nm")
}

MuEtOH_nest |>
  unnest(cols = c(LinearPredict)) |>
  ggplot() +
  geom_point(aes(x = EthanolConcentration, y = Mumax_hour)) +
  geom_line(aes(x = EthanolConcentration , y = .fitted)) +
  #geom_point(aes(x = Ethanol (v/v %), y = .resid), colour = "red") +
  facet_grid(cols = vars(Source, BrothType, GrowthTempC), rows = vars(nm), labeller = labeller(GrowthTempC = temp_labeller , nm = nm_labeller)) +
   labs(x = "Ethanol Concentration % (v/v)") +
  theme_bw()
```


```{r show mumax vs EtOH fit parameters}
MuEtOH_nest |>
unnest(cols = c(LinearTidy)) |>
 select(-c(data, LinearFit, LinearParam, LinearPredict)) |>
  select(-c(statistic)) |>
  pivot_wider(id_cols = c(SampleID, nm), names_from = term, values_from = c(estimate, std.error, p.value)) |>
  kable()
```




